### 2020-06-22 documentation/paper/polymer.tex

%!TEX root=main.tex

\subsection{Polymer}
Installing Polymer is straight forward.