### 2020-08-10 documentation/paper/testingMethods.tex
some layouting in main, begin testing methods and some work on giraph setup guide

%!TEX root=main.tex
\subsection{Hardware and Software}
For testing all systems we used 5 machines with 96 CPU cores each (48 physical) and 256 GB of RAM. One of those machines only had 128 GB of RAM, this one was only used as part of the calculation cluster for the distributed systems.
All servers were running Ubuntu 19.x.
Setup of each framework was performed according to our provided installation guides.

\subsection{Benchmark setup}
We measured the total runtime of each process as well as the actual calculation times. The console output of each frameworks was used as an indicator for when loading the graph data was finished. This proved to be more difficult than expected in some cases.

Each test case consisting of graph, framework and in some cases even the algorithm was run 10 times, allowing us to smooth slight variations in the measured times. This also helps to reduce the error we introduced by measuring time through the console log.

All benchmarks were initiated by our benchmark script that is available in our repository.

We will compare both the computation times as well es the execution times. This way we compare how much setup time each framework has.