### 2020-08-11 documentation/paper/sections/overview/galois.tex
improved galois writeup

%!TEX root=../../main.tex

Galois\cite{Galois} is a general purpose library designed for parallel programming. The Galois system supports fine grain tasks, allows for autonomous, speculative execution of these tasks and grants control over the task scheduling policies to the application. It also simplifies the implementation of parallel applications by providing an implicitly parallel unordered-set iterators.

For graph analytics purposes a topology aware work stealing scheduler, a priority scheduler and a library of scalable data structures have been implemented. Galois includes applications for many graph analytics problems, among these are single-source shorthest-paths (sssp), breath-first-search (bfs) and pagerank. For most of these applications Galois offers several different algorithms to perform these analytics problems and many setting options like the amount of threads used or policies for splitting the graph. All of these applications can be executed in shared memory systems and, due to the Gluon integration, with a few modifications in a distributed environment.

Gluon\cite{vertGalois} is a framework written by the Galois team as a middleware to write graph analysis applications for distributed systems. It reduces the communication overhead needed in distributed environments by exploiting structural and temporal invariants.

The code of Gluon is embedded in Galois. It is possible to integrate Gluon in other frameworks too, which the Galois team showed in their paper\cite{vertGalois}.