### 2020-08-15 documentation/paper/sections/overview/polymer.tex
polymer overview

%!TEX root=../../main.tex

Polymer is very similar to Ligra, in fact Polymer inherits the programming interfaces \texttt{EdgeMap} and \texttt{VertexMap} from Ligra as its main interface.

Polymer is a vertex-centric framework, that tries to circumvent some of the random memory access drawbacks of such a design. It treats a NUMA machine as a distributed cluster and splits work and graph data accordingly between the nodes.\cite[p.187]{Polymer}
Application-defined data is not distributed. Other runtime state data is allocated in a distributed way but only accessed through a global lookup table.\cite[p.184]{Polymer}