### 2020-08-22 documentation/paper/sections/overview/gemini.tex

%!TEX root=../../main.tex

Gemini is a framework for parallel graph processing\cite{Gemini}.
In comparison to the other NUMA-aware frameworks we discuss here, Gemini does not support shared memory parallel processing.
Computation is entirely message based using MPI.
Gemini is lightweight and only provides basic functionality.
A basic API hides data and computation distribution details from the user writing applications.
There are five algorithms (which ones?) already implemented.