### 2020-08-28 documentation/paper/main.tex
erste fixes der Rückmeldung

\documentclass[conference,a4paper]{IEEEtran}

\usepackage[noadjust]{cite}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage{textcomp}
\usepackage{pifont}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{amssymb}

\usepackage{tabularx}
\def\tabularxcolumn#1{m{#1}} % vertical centering in tabularx cells
\newcolumntype{Y}{>{\centering\arraybackslash}X} % cell type Y for centered text with automatic line breaks
\newcommand*\rot{\rotatebox{90}} % rotates content 90deg ccw, mainly for table headers
\newcommand*\OK{\ding{51}} % a checkmark-style sign

\bibliography{literature.bib}

\title{A Comparison of Graph Processing Systems}

\author{\IEEEauthorblockN{Simon König}
\IEEEauthorblockA{(3344789)\\
st156571@stud.uni-stuttgart.de}
\and
\IEEEauthorblockN{Leon Matzner}
\IEEEauthorblockA{(3315161)\\
st155698@stud.uni-stuttgart.de}
\and
\IEEEauthorblockN{Felix Rollbühler}
\IEEEauthorblockA{(3310069)\\
st154960@stud.uni-stuttgart.de}
\and
\IEEEauthorblockN{Jakob Schmid}
\IEEEauthorblockA{(3341630)\\
st157100@stud.uni-stuttgart.de}}

\date{\today}

\begin{document}

\maketitle

\begin{abstract}
TODO
\end{abstract}

\begin{IEEEkeywords}
graphs, distributed computing, Galois, Ligra, Polymer, Giraph, Gluon, Gemini
\end{IEEEkeywords}

\section{Introduction}
In recent years graph sizes have increased significantly [?] .
Thus performance and memory efficiency of the graph analysis applications is now more important than ever.

Many applications like

Wachsende graphen,
benötigt für folgende Applikationen,
verwenden graph algorithmen,
diese müssen schnell und verteilt laufen
dafür wurden Frameworks entwickelt...

und das noch schön formulieren

We compare several non-uniform memory access (NUMA) aware systems in terms of their performance on three graph algorithms (PageRank, SSSP, BFS).
The comparison is performed on both real world data sets and synthetic graphs.

To provide a comparison to a non-NUMA aware system, Giraph\cite{Giraph} is included in the testing. Giraph itself has often been compared to other state of the art systems like Pregel or GraphX.

%One of the most famous is the (closed source) distributed graph processing system Pregel. Pregel and several open source versions of .
%This results in suboptimal performance, due to missing cache locality, because the Pregel like systems are written in an object oriented manner relying on pointer chasing mechanisms.

To this end several non-uniform memory access (NUMA) aware systems were proposed like Polymer \cite{Polymer}, Galois \cite{Galois} or Ligra \cite{Ligra}.

This paper makes the following contributions:
\begin{itemize}
  \item Comparison of several state-of-the-art graph processing frameworks
  \item
  \item
  \item
\end{itemize}

\section{Preliminaries}

\paragraph{Graphs and Paths}
An \emph{unweighted graph} is the pair $G=(V,E)$ where the \emph{vertex set} is $V\subseteq\mathbb N$ and the $E$ is the \emph{edge set}.
The edge set describes a number of connections or relations between two vertices. Depending on these relations, a graph can be directed or undirected. For a \emph{directed graph} the edge set becomes 
\begin{equation*}
  E\subseteq\{(x,y)\,|\, x,y\in V, x\neq y\}
\end{equation*}
and in the \emph{undirected} case $E$ is a set of two-sets 
\begin{equation*}
  E\subseteq\{\{x,y\}\,|\, x,y\in V, x\neq y\}.
\end{equation*}

The main difference is thus, that in the case of a directed graph a connection between $s$ and $t$ is not the same as a connection between $t$ and $s$ -- the direction matters.

Independently of the graph being directed or not, a graph can be \emph{weighted}. In this case the edge set is expanded by a numerical value $E_{\text{weighted}}=E_{\text{unweighted}}\times \mathbb R$ further describing the relation.

The size of a graph is often described in the number of edges $|E|$ because this number is typically much larger than the number of vertices $|V|$.

A \emph{Path} from start $s$ to target $t$ is a set of edges $P\subseteq E^n$ with
\begin{equation*}
   P=((x_0,x_1),(x_1,x_2),\ldots, (x_{n-1},x_n))
\end{equation*}
where all $(x_i,x_{i+1})\in E$ and $x_0=s, x_n=t$.

Thus we call a target $t$ \emph{accessible} from $s$ if a Path from $s$ to $t$ exists.

\paragraph{Single-Source Shortest-Paths}
Single-Source Shortest-Paths (SSSP) describes the problem of finding a path along some edges of the input graph from a given start node to a target.

Input to the problem is a weighted graph $G=(V,E)$ and a start node $s\in V$. Output is the shortest possible distance from $s$ to each node in $V$. 
The distance is defined as the sum of edge weights $w_i$ on a path from $s$ to the target.
In the case of a unweighted graph, the distance is often described in \emph{hops}, i.e. the number of edges on a path.

The most common implementations are Dijkstra's algorithm or BellmanFord.

\paragraph{Breadth-First Search}
Breadth-first search (BFS) is a search problem on a graph. It usually requires an unweighted graph and a start vertex as input. In some cases a target vertex is also given.

The output is usually a set of vertices that are accessible from the start vertex. In the case of a target being given, the output is true if a path from start to target exists or false otherwise.

\paragraph{PageRank}

\paragraph{Pregel Model}

% Google, to overcome, these challenges came up with Pregel.
% Provides scalability
% Fault-tolerance
% Flexibility to express arbitrary algorithms
% The high level organization of Pregel programs
% is inspired by Valiant’s Bulk Synchronous
%  Parallel model 
\paragraph{Congest Model}

\section{Overview of the framworks}
\subsection{Galois and Gluon}
\input{sections/overview/galois}

\subsection{Gemini}
\input{sections/overview/gemini}

\subsection{Giraph}
\input{sections/overview/giraph}

\subsection{Ligra}
\input{sections/overview/ligra}

\subsection{Polymer}
\input{sections/overview/polymer}

\section{And}
\input{sections/graphFormats}%hier oder supplementary data?

\section{Testing methods}
\input{sections/testingMethods}

\section{Results}
\input{sections/results/results}

\section{Discussion}

\section{Conclusion}
The conclusion goes here.

\clearpage

\section*{Acknowledgment}
We are using the graph frameworks Galois~\cite{Galois}, Ligra~\cite{Ligra}, Polymer~\cite{Polymer}, Gemini~\cite{Gemini} as well as Apache Giraph~\cite{Giraph}.

Also we use Gluon~\cite{vertGalois} for the distributed Galois setups.

Gemini~\cite{Gemini}

%We would like to thank our supervisor Heiko Geppert for continued guidance and support.

\section*{Supplementary Data}\label{supplementaryData}
We have written a number of conversion tools and installation guides to help
users or developers with the use of the tested frameworks.

Our GitHub repository: \url{http://www.github.com/serengti/Forschungsprojekt}.

% Beim Hinzufügen neuer Referenzen
% 1. die beiden Zeilen nach diesem Kommentar auskommentieren.
%    (gemeint ist \bibliographystyle... und \bibliography..)
%    Dann alle cache files im Ordner löschen.
% 2. Die folgenden Commands ausführen
% pdflatex main.tex
% bibtex main
% pdflatex main.tex
%    das letzte wirft dann Fehler aber das stimmt so..
% 3. Jetzt main.bbl öffnen und den Inhalt hier drunter rein kopieren.
%    Man kann kein \input benutzen, weil main.bbl leer ist, wenn die beiden
%    Zeilen auskommentiert sind. Aber wenn sie nicht auskommentiert sind
%    kann kein PDF erzeugt werden.. Was zur Hölle denken die sich.
% 4. Die folgenden beiden Zeilen auskommentieren und pdflatex
%    ausführen wie gewohnt.

% \bibliographystyle{IEEEtran}
% \bibliography{literature}

% Alles ab hier kommt aus main.bbl

% Generated by IEEEtran.bst, version: 1.12 (2007/01/11)
% Generated by IEEEtran.bst, version: 1.12 (2007/01/11)
% Generated by IEEEtran.bst, version: 1.12 (2007/01/11)
\begin{thebibliography}{1}
\providecommand{\url}[1]{#1}
\csname url@samestyle\endcsname
\providecommand{\newblock}{\relax}
\providecommand{\bibinfo}[2]{#2}
\providecommand{\BIBentrySTDinterwordspacing}{\spaceskip=0pt\relax}
\providecommand{\BIBentryALTinterwordstretchfactor}{4}
\providecommand{\BIBentryALTinterwordspacing}{\spaceskip=\fontdimen2\font plus
\BIBentryALTinterwordstretchfactor\fontdimen3\font minus
  \fontdimen4\font\relax}
\providecommand{\BIBforeignlanguage}[2]{{%
\expandafter\ifx\csname l@#1\endcsname\relax
\typeout{** WARNING: IEEEtran.bst: No hyphenation pattern has been}%
\typeout{** loaded for the language `#1'. Using the pattern for}%
\typeout{** the default language instead.}%
\else
\language=\csname l@#1\endcsname
\fi
#2}}
\providecommand{\BIBdecl}{\relax}
\BIBdecl

\bibitem{Galois}
\BIBentryALTinterwordspacing
D.~Nguyen, A.~Lenharth, and K.~Pingali, ``A lightweight infrastructure for
  graph analytics,'' in \emph{Proceedings of the Twenty-Fourth ACM Symposium on
  Operating Systems Principles}, ser. SOSP ’13.\hskip 1em plus 0.5em minus
  0.4em\relax New York, NY, USA: Association for Computing Machinery, 2013, p.
  456–471. [Online]. Available: \url{https://doi.org/10.1145/2517349.2522739}
\BIBentrySTDinterwordspacing

\bibitem{vertGalois}
\BIBentryALTinterwordspacing
R.~Dathathri, G.~Gill, L.~Hoang, H.-V. Dang, A.~Brooks, N.~Dryden, M.~Snir, and
  K.~Pingali, ``Gluon: A communication-optimizing substrate for distributed
  heterogeneous graph analytics,'' in \emph{Proceedings of the 39th ACM SIGPLAN
  Conference on Programming Language Design and Implementation}, ser. PLDI
  2018.\hskip 1em plus 0.5em minus 0.4em\relax New York, NY, USA: Association
  for Computing Machinery, 2018, p. 752–768. [Online]. Available:
  \url{https://doi.org/10.1145/3192366.3192404}
\BIBentrySTDinterwordspacing

\bibitem{Polymer}
\BIBentryALTinterwordspacing
K.~Zhang, R.~Chen, and H.~Chen, ``Numa-aware graph-structured analytics,'' in
  \emph{Proceedings of the 20th ACM SIGPLAN Symposium on Principles and
  Practice of Parallel Programming}, ser. PPoPP 2015.\hskip 1em plus 0.5em
  minus 0.4em\relax New York, NY, USA: Association for Computing Machinery,
  2015, p. 183–193. [Online]. Available:
  \url{https://doi.org/10.1145/2688500.2688507}
\BIBentrySTDinterwordspacing

\bibitem{adjListFormat}
\BIBentryALTinterwordspacing
J.~Shun, G.~Blelloch, J.~Fineman, P.~Gibbons, A.~Kyrola, K.~Tangwonsan, and
  H.~V. Simhadri. (2020, Jun.) {Problem Based Benchmark Suite}. graphIO.html.
  [Online]. Available: \url{http://www.cs.cmu.edu/~pbbs/benchmarks/}
\BIBentrySTDinterwordspacing

\bibitem{Ligra}
\BIBentryALTinterwordspacing
J.~Shun and G.~E. Blelloch, ``Ligra: A lightweight graph processing framework
  for shared memory,'' in \emph{Proceedings of the 18th ACM SIGPLAN Symposium
  on Principles and Practice of Parallel Programming}, ser. PPoPP ’13.\hskip
  1em plus 0.5em minus 0.4em\relax New York, NY, USA: Association for Computing
  Machinery, 2013, p. 135–146. [Online]. Available:
  \url{https://doi.org/10.1145/2442516.2442530}
\BIBentrySTDinterwordspacing

\bibitem{Gemini}
\BIBentryALTinterwordspacing
X.~Zhu, W.~Chen, W.~Zheng, and X.~Ma, ``Gemini: A computation-centric
  distributed graph processing system,'' in \emph{12th {USENIX} Symposium on
  Operating Systems Design and Implementation ({OSDI} 16)}.\hskip 1em plus
  0.5em minus 0.4em\relax Savannah, GA: {USENIX} Association, Nov. 2016, pp.
  301--316. [Online]. Available:
  \url{https://www.usenix.org/conference/osdi16/technical-sessions/presentation/zhu}
\BIBentrySTDinterwordspacing

\bibitem{Giraph}
\BIBentryALTinterwordspacing
A.~S. Foundation. (2020, Jun.) {Apache Giraph}. [Online]. Available:
  \url{https://giraph.apache.org}
\BIBentrySTDinterwordspacing

\bibitem{konect}
J.~Kunegis, ``Konect: the koblenz network collection,'' 05 2013, pp.
  1343--1350.

\end{thebibliography}

% Alles bis hier kommt aus main.bbl

\end{document}