### 2020-09-03 documentation/paper/sections/overview/ligra.tex

%!TEX root=../../main.tex

Ligra is a lightweight graph processing framework for shared memory machines \cite{Ligra}. It offers a vertex-centric programming interface which can be used to apply a function to each vertex or outgoing edge of a set of vertices in parallel. While doing so the framework can generate a set of active vertices for the next iteration. This abstraction makes it well suited for writing graph traversal algorithms.

When mapping over edges Ligra optimizes algorithms by switching between a push-based and a pull-based approach based on the size of the set. When the size of the set is above a threshold a pull approach is used. If the threshold is not reached a push approach is used.