### 2020-09-04 documentation/paper/sections/overview/overview.tex

%!TEX root=../../main.tex

\section{Overview}
In this section we describe the functionality and characteristics of several state-of-the-art graph processing frameworks. 
We start with a short overview describing functionality and characteristics of each framework, followed by their required input data formats.

Further

//
\subsection{Galois and Gluon}
\input{sections/overview/galois}

\subsection{Gemini}
\input{sections/overview/gemini}

\subsection{Giraph}
\input{sections/overview/giraph}

\subsection{Ligra}
\input{sections/overview/ligra}

\subsection{Polymer}
\input{sections/overview/polymer}

\input{sections/graphFormats}