### 2020-09-15 documentation/paper/main.tex
new references

\documentclass[a4paper]{IEEEtran}

\usepackage[noadjust]{cite}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage{textcomp}
\usepackage{pifont}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{amssymb}

\usepackage{tabularx}
\def\tabularxcolumn#1{m{#1}} % vertical centering in tabularx cells
\newcolumntype{Y}{>{\centering\arraybackslash}X} % cell type Y for centered text with automatic line breaks
\newcommand*\rot{\rotatebox{45}} % rotates content 90deg ccw, mainly for table headers
\newcommand*\OK{\ding{51}} % a checkmark-style sign

\newcommand{\todo}[1]{{\color{red!90!black}//\bf{TODO:} #1}}

\bibliography{literature.bib}

\title{A Comparison of Graph Processing Systems}

\author{\IEEEauthorblockN{Simon König}
\IEEEauthorblockA{(3344789)
st156571@stud.uni-stuttgart.de}

\and

\IEEEauthorblockN{Leon Matzner}
\IEEEauthorblockA{(3315161)
st155698@stud.uni-stuttgart.de}

\and

\IEEEauthorblockN{Felix Rollbühler}
\IEEEauthorblockA{(3310069)
st154960@stud.uni-stuttgart.de}

\and

\IEEEauthorblockN{Jakob Schmid}
\IEEEauthorblockA{(3341630)
st157100@stud.uni-stuttgart.de}}

\date{\today}

\begin{document}

\maketitle

\begin{abstract}
\todo{Section}
\end{abstract}

\begin{IEEEkeywords}
graphs, distributed computing, Galois, Ligra, Polymer, Giraph, Gluon, Gemini
\end{IEEEkeywords}

\section*{TODOs}
\begin{itemize}
  \item Appendix für Setup Guides - md zu pdf
\end{itemize}

\section{Introduction}
\todo{Section}
% In recent years graph sizes have increased significantly [?] .
% Thus performance and memory efficiency of the graph analysis applications is now more important than ever.

% Many applications like

% Wachsende graphen,
% benötigt für folgende Applikationen,
% verwenden graph algorithmen,
% diese müssen schnell und verteilt laufen
% dafür wurden Frameworks entwickelt...

% und das noch schön formulieren

% We compare several non-uniform memory access (NUMA) aware systems in terms of their performance on three graph algorithms (PageRank, SSSP, BFS).
% The comparison is performed on both real world data sets and synthetic graphs.

% To provide a comparison to a non-NUMA aware system, Giraph\cite{Giraph} is included in the testing. Giraph itself has often been compared to other state of the art systems like Pregel or GraphX.

%One of the most famous is the (closed source) distributed graph processing system Pregel. Pregel and several open source versions of .
%This results in suboptimal performance, due to missing cache locality, because the Pregel like systems are written in an object oriented manner relying on pointer chasing mechanisms.

% To this end several non-uniform memory access (NUMA) aware systems were proposed like Polymer \cite{Polymer}, Galois \cite{Galois} or Ligra \cite{Ligra}.

This paper makes the following contributions:
\begin{itemize}
  \item Comparison of several state-of-the-art graph processing frameworks
  \item
  \item
  \item
\end{itemize}

\input{sections/relatedWork}

\input{sections/preliminiaries}

\input{sections/overview/overview}

\input{sections/testingMethods}

\input{sections/results/results}

\section{Discussion}
\todo{Section}

\section{Conclusion}
\todo{Section}

\clearpage

\section*{Acknowledgments}
%We would like to thank our supervisor Heiko Geppert for his continued guidance and support.

\appendices
\todo{Beide zusammenfassen}
\section{Installation Guides}\label{app:installationGuides}
The complete installation guides with steps on how to run the same algorithms we used is available in our repository\footnote{\url{https://github.com/SerenGTI/Forschungsprojekt/tree/master/documentation}}

In case of frameworks like Giraph, where for example no BFS algorithm is provided by the framework itself, the algorithm and a guide on how to include it in the framework can be found in the repository as well.

\section{Conversion Tool}\label{app:conversionTool}
In order to

We have written a number of conversion tools and installation guides to help
users or developers with the use of the tested frameworks.

Our GitHub repository: \url{http://www.github.com/serengti/Forschungsprojekt}.

% Beim Hinzufügen neuer Referenzen
% 1. die beiden Zeilen nach diesem Kommentar auskommentieren.
%    (gemeint ist \bibliographystyle... und \bibliography..)
%    Dann alle cache files im Ordner löschen.
% 2. Die folgenden Commands ausführen
% pdflatex main.tex
% bibtex main
% pdflatex main.tex
%    das letzte wirft dann Fehler aber das stimmt so..
% 3. Jetzt main.bbl öffnen und den Inhalt hier drunter rein kopieren.
%    Man kann kein \input benutzen, weil main.bbl leer ist, wenn die beiden
%    Zeilen auskommentiert sind. Aber wenn sie nicht auskommentiert sind
%    kann kein PDF erzeugt werden.. Was zur Hölle denken die sich.
% 4. Die folgenden beiden Zeilen auskommentieren und pdflatex
%    ausführen wie gewohnt.

%\bibliographystyle{IEEEtran}
%\bibliography{literature}

% Alles ab hier kommt aus main.bbl

% Generated by IEEEtran.bst, version: 1.12 (2007/01/11)
\begin{thebibliography}{10}
\providecommand{\url}[1]{#1}
\csname url@samestyle\endcsname
\providecommand{\newblock}{\relax}
\providecommand{\bibinfo}[2]{#2}
\providecommand{\BIBentrySTDinterwordspacing}{\spaceskip=0pt\relax}
\providecommand{\BIBentryALTinterwordstretchfactor}{4}
\providecommand{\BIBentryALTinterwordspacing}{\spaceskip=\fontdimen2\font plus
\BIBentryALTinterwordstretchfactor\fontdimen3\font minus
  \fontdimen4\font\relax}
\providecommand{\BIBforeignlanguage}[2]{{%
\expandafter\ifx\csname l@#1\endcsname\relax
\typeout{** WARNING: IEEEtran.bst: No hyphenation pattern has been}%
\typeout{** loaded for the language `#1'. Using the pattern for}%
\typeout{** the default language instead.}%
\else
\language=\csname l@#1\endcsname
\fi
#2}}
\providecommand{\BIBdecl}{\relax}
\BIBdecl

\bibitem{newman2010networks}
\BIBentryALTinterwordspacing
M.~Newman, \emph{Networks: An Introduction}.\hskip 1em plus 0.5em minus
  0.4em\relax OUP Oxford, 2010. [Online]. Available:
  \url{https://books.google.de/books?id=-DgTDAAAQBAJ}
\BIBentrySTDinterwordspacing

\bibitem{Polymer}
\BIBentryALTinterwordspacing
K.~Zhang, R.~Chen, and H.~Chen, ``Numa-aware graph-structured analytics,'' in
  \emph{Proceedings of the 20th ACM SIGPLAN Symposium on Principles and
  Practice of Parallel Programming}, ser. PPoPP 2015.\hskip 1em plus 0.5em
  minus 0.4em\relax New York, NY, USA: Association for Computing Machinery,
  2015, p. 183–193. [Online]. Available:
  \url{https://doi.org/10.1145/2688500.2688507}
\BIBentrySTDinterwordspacing

\bibitem{Ligra}
\BIBentryALTinterwordspacing
J.~Shun and G.~E. Blelloch, ``Ligra: A lightweight graph processing framework
  for shared memory,'' in \emph{Proceedings of the 18th ACM SIGPLAN Symposium
  on Principles and Practice of Parallel Programming}, ser. PPoPP ’13.\hskip
  1em plus 0.5em minus 0.4em\relax New York, NY, USA: Association for Computing
  Machinery, 2013, p. 135–146. [Online]. Available:
  \url{https://doi.org/10.1145/2442516.2442530}
\BIBentrySTDinterwordspacing

\bibitem{pregel}
\BIBentryALTinterwordspacing
G.~Malewicz, M.~H. Austern, A.~J. Bik, J.~C. Dehnert, I.~Horn, N.~Leiser, and
  G.~Czajkowski, ``Pregel: A system for large-scale graph processing,'' in
  \emph{Proceedings of the 2010 ACM SIGMOD International Conference on
  Management of Data}, ser. SIGMOD '10.\hskip 1em plus 0.5em minus 0.4em\relax
  New York, NY, USA: Association for Computing Machinery, 2010, p. 135–146.
  [Online]. Available: \url{https://doi.org/10.1145/1807167.1807184}
\BIBentrySTDinterwordspacing

\bibitem{pagerank}
\BIBentryALTinterwordspacing
S.~Brin and L.~Page, ``The anatomy of a large-scale hypertextual web search
  engine,'' \emph{Computer Networks and ISDN Systems}, vol.~30, no.~1, pp. 107
  -- 117, 1998, proceedings of the Seventh International World Wide Web
  Conference. [Online]. Available:
  \url{http://www.sciencedirect.com/science/article/pii/S016975529800110X}
\BIBentrySTDinterwordspacing

\bibitem{tao}
\BIBentryALTinterwordspacing
K.~Pingali, D.~Nguyen, M.~Kulkarni, M.~Burtscher, M.~A. Hassaan, R.~Kaleem,
  T.-H. Lee, A.~Lenharth, R.~Manevich, M.~M\'{e}ndez-Lojo, D.~Prountzos, and
  X.~Sui, ``The tao of parallelism in algorithms,'' in \emph{Proceedings of the
  32nd ACM SIGPLAN Conference on Programming Language Design and
  Implementation}, ser. PLDI '11.\hskip 1em plus 0.5em minus 0.4em\relax New
  York, NY, USA: Association for Computing Machinery, 2011, p. 12–25.
  [Online]. Available: \url{https://doi.org/10.1145/1993498.1993501}
\BIBentrySTDinterwordspacing

\bibitem{bsp}
\BIBentryALTinterwordspacing
L.~G. Valiant, ``A bridging model for parallel computation,'' \emph{Commun.
  ACM}, vol.~33, no.~8, p. 103–111, Aug. 1990. [Online]. Available:
  \url{https://doi.org/10.1145/79173.79181}
\BIBentrySTDinterwordspacing

\bibitem{Giraph}
\BIBentryALTinterwordspacing
{Apache Software Foundation}. (2020, Jun.) {Apache Giraph}. [Online].
  Available: \url{https://giraph.apache.org}
\BIBentrySTDinterwordspacing

\bibitem{graphx}
J.~E. Gonzalez, R.~S. Xin, A.~Dave, D.~Crankshaw, M.~J. Franklin, and
  I.~Stoica, ``Graphx: Graph processing in a distributed dataflow framework,''
  in \emph{Proceedings of the 11th USENIX Conference on Operating Systems
  Design and Implementation}, ser. OSDI'14.\hskip 1em plus 0.5em minus
  0.4em\relax USA: USENIX Association, 2014, p. 599–613.

\bibitem{powergraph}
\BIBentryALTinterwordspacing
J.~E. Gonzalez, Y.~Low, H.~Gu, D.~Bickson, and C.~Guestrin, ``Powergraph:
  Distributed graph-parallel computation on natural graphs,'' in \emph{10th
  {USENIX} Symposium on Operating Systems Design and Implementation ({OSDI}
  12)}.\hskip 1em plus 0.5em minus 0.4em\relax Hollywood, CA: {USENIX}
  Association, Oct. 2012, pp. 17--30. [Online]. Available:
  \url{https://www.usenix.org/conference/osdi12/technical-sessions/presentation/gonzalez}
\BIBentrySTDinterwordspacing

\bibitem{mapreduce}
J.~Dean and S.~Ghemawat, ``Mapreduce: Simplified data processing on large
  clusters,'' in \emph{In Proceedings of Operating Systems Design and
  Implementation (OSDI}, pp. 137--150.

\bibitem{hugepages}
\BIBentryALTinterwordspacing
A.~Panwar, A.~Prasad, and K.~Gopinath, ``Making huge pages actually useful,''
  in \emph{Proceedings of the Twenty-Third International Conference on
  Architectural Support for Programming Languages and Operating Systems}, ser.
  ASPLOS '18.\hskip 1em plus 0.5em minus 0.4em\relax New York, NY, USA:
  Association for Computing Machinery, 2018, p. 679–692. [Online]. Available:
  \url{https://doi.org/10.1145/3173162.3173203}
\BIBentrySTDinterwordspacing

\bibitem{origGalois}
\BIBentryALTinterwordspacing
M.~Kulkarni, K.~Pingali, B.~Walter, G.~Ramanarayanan, K.~Bala, and L.~P. Chew,
  ``Optimistic parallelism requires abstractions,'' in \emph{Proceedings of the
  28th ACM SIGPLAN Conference on Programming Language Design and
  Implementation}, ser. PLDI '07.\hskip 1em plus 0.5em minus 0.4em\relax New
  York, NY, USA: Association for Computing Machinery, 2007, p. 211–222.
  [Online]. Available: \url{https://doi.org/10.1145/1250734.1250759}
\BIBentrySTDinterwordspacing

\bibitem{Galois}
\BIBentryALTinterwordspacing
D.~Nguyen, A.~Lenharth, and K.~Pingali, ``A lightweight infrastructure for
  graph analytics,'' in \emph{Proceedings of the Twenty-Fourth ACM Symposium on
  Operating Systems Principles}, ser. SOSP ’13.\hskip 1em plus 0.5em minus
  0.4em\relax New York, NY, USA: Association for Computing Machinery, 2013, p.
  456–471. [Online]. Available: \url{https://doi.org/10.1145/2517349.2522739}
\BIBentrySTDinterwordspacing

\bibitem{vertGalois}
\BIBentryALTinterwordspacing
R.~Dathathri, G.~Gill, L.~Hoang, H.-V. Dang, A.~Brooks, N.~Dryden, M.~Snir, and
  K.~Pingali, ``Gluon: A communication-optimizing substrate for distributed
  heterogeneous graph analytics,'' in \emph{Proceedings of the 39th ACM SIGPLAN
  Conference on Programming Language Design and Implementation}, ser. PLDI
  2018.\hskip 1em plus 0.5em minus 0.4em\relax New York, NY, USA: Association
  for Computing Machinery, 2018, p. 752–768. [Online]. Available:
  \url{https://doi.org/10.1145/3192366.3192404}
\BIBentrySTDinterwordspacing

\bibitem{Gemini}
\BIBentryALTinterwordspacing
X.~Zhu, W.~Chen, W.~Zheng, and X.~Ma, ``Gemini: A computation-centric
  distributed graph processing system,'' in \emph{12th {USENIX} Symposium on
  Operating Systems Design and Implementation ({OSDI} 16)}.\hskip 1em plus
  0.5em minus 0.4em\relax Savannah, GA: {USENIX} Association, Nov. 2016, pp.
  301--316. [Online]. Available:
  \url{https://www.usenix.org/conference/osdi16/technical-sessions/presentation/zhu}
\BIBentrySTDinterwordspacing

\bibitem{konect}
J.~Kunegis, ``Konect: the koblenz network collection,'' 05 2013, pp.
  1343--1350.

\bibitem{snap}
\BIBentryALTinterwordspacing
{Stanford University}. (2020, Sep.) Stanford network analysis project.
  [Online]. Available: \url{https://snap.stanford.edu}
\BIBentrySTDinterwordspacing

\bibitem{pbbs}
\BIBentryALTinterwordspacing
J.~Shun, G.~Blelloch, J.~Fineman, P.~Gibbons, A.~Kyrola, K.~Tangwonsan, and
  H.~V. Simhadri. (2020, Aug.) {Problem Based Benchmark Suite }. [Online].
  Available: \url{http://www.cs.cmu.edu/~pbbs/}
\BIBentrySTDinterwordspacing

\end{thebibliography}

\end{document}