### 2020-09-19 documentation/paper/sections/results/results.tex

%!TEX root=../../main.tex

\subsection{Results}
\todo{Einleitung}

\subsubsection{Lessons Learned}

\todo{umbenennen?}

We would like to raise some issues we encountered first while installing and configuring and second while running the different frameworks.

\begin{enumerate}
	\item During setup and benchmark of Gemini, we encountered several bugs in the cloned repository. These include non zero-terminated strings or even missing return statements.

	The errors rendered the code as-is unable to perform calculations, forcing us to fork the repository and modify the source code. Our changes can be found in one of our repositories\footnote{\url{https://github.com/jasc7636/GeminiGraph}}.

	\item Furthermore, we would like to address the setup of Hadoop for Giraph. It requires multiple edits in \texttt{xml} files that aren't easily automized. This makes the setup rather time consuming, especially if reconfiguration is needed later on.
	\item In order for Giraph to run, several Java tasks (the Hadoop infrastructure) have to be constantly running in the background. While we don't expect this to have a significant performance impact on other tasks, it is still suboptimal.
	\item Giraph ran us into disk space problems on multiple occasions. First, deleting files on the Hadoop distributed file system (HDFS) does not immediately free up disk space because the files are moved to a \emph{recycling bin}-like location. Second, some log files that can easily be multiple gigabytes in size are stored outside of the HDFS and are never mentioned in the Giraph documentation.
\end{enumerate}

On a plus side, setup of the frameworks Polymer and Ligra was straight forward and did not require any special treatment.

\input{sections/results/sssp}

\input{sections/results/bfs}

\input{sections/results/pr}

\subsubsection{Comparison of a Single-Node to the Distributed Cluster}
\begin{table}
	\caption{Execution Times on a SingleNode vs. 5-Node Distributed Cluster}
	\label{tbl:SSSP_execTimeComparison}
	\renewcommand{\arraystretch}{1.2}
	\centering
	\begin{tabular}{ccr@{\tabskip 1 \tabcolsep}r
	r@{\tabskip 1 \tabcolsep}r
	r@{\tabskip 1 \tabcolsep}r}
		\toprule
		&&\multicolumn{2}{c}{\bf Galois}&\multicolumn{2}{c}{\bf Gemini}&\multicolumn{2}{c}{\bf Giraph}\\
		\cmidrule{3-4}\cmidrule{5-6}\cmidrule{7-8}
		&\bf Graph&1N&5N&1N&5N&1N&5N\\
		\midrule
		\multirow{7}{0.5ex}{\rotatebox{90}{\bf SSSP}}&flickr & 0.3 & 2.4 & 0.6 & 2.5 & 38.8 & 42.6 \\
		& orkut & 0.8 & 4.2 & 13.9 & 17.3 & 135.9 & 60.4 \\
		& wikipedia & 1.8 & 14.2 & 48.5 & 74.5 & 377.2 & 100.8 \\
		& twitter & 10.8 & 40.0 & 383.0 & 391.1 & - & 349.9 \\
		& rMat27 & 16.0 & 39.0 & 563.5 & 386.0 & - & 565.8 \\
		& friendster & 14.4 & 58.5 & 742.6 & 568.9 & - & 444.0 \\
		& rMat28 & 27.8 & 71.5 & 1236.7 & 792.0 & - & 1180.2 \\
		\midrule
		\multirow{7}{0.5ex}{\rotatebox{90}{\bf BFS}}& flickr & 0.6 & 2.3 & 0.7 & 2.4 & 38.9 & 41.9 \\
		& orkut & 0.9 & 4.0 & 12.5 & 13.8 & 128.2 & 58.1 \\
		& wikipedia & 2.4 & 13.2 & 44.9 & 60.9 & 360.7 & 95.1 \\
		& twitter & 14.2 & 37.9 & 355.0 & 293.2 & - & 322.3 \\
		& rMat27 & 14.6 & 34.4 & 540.8 & 305.3 & - & 539.2 \\
		& friendster & 12.8 & 47.6 & 708.4 & 450.2 & - & 412.3 \\
		& rMat28 & 33.1 & 67.1 & 1178.7 & 634.3 & - & 1135.6 \\
		\midrule
		\multirow{7}{0.5ex}{\rotatebox{90}{\bf PR}}& flickr & 0.3\txtdagger & 2.6 & 0.7 & 2.2 & 45.6 & 44.1 \\
		& orkut & 0.7\txtdagger & 8.0 & 13.1 & 19.4 & 561.3 & 106.2 \\
		& wikipedia & 1.7\txtdagger & 206.5 & 44.4 & 54.4 & 1501.3 & 252.6 \\
		& twitter & 8.7\txtdagger & 910.9 & 359.9 & 363.5 & - & 1200.2 \\
		& rMat27 & 19.2\txtdagger & 1287.4 & 536.3 & 376.3 & - & 1369.5 \\
		& friendster & 20.4\txtdagger & 593.3 & 716.5 & 591.1 & - & 1655.0 \\
		& rMat28 & 46.0\txtdagger & 2540.2 & 1188.5 & 774.2 & - & - \\
		\bottomrule
		\multicolumn{8}{l}{(\txtdagger) Results of Galois Pull shown.}
	\end{tabular}
\end{table}
Until now, we have only compared the frameworks with each other, under the same setup circumstances. Hence, a comparison of the same framework, in single-node versus distributed setup is the content of this section.
Of course only frameworks that were tested in both setups are shown.
We focus primarily on the results of the execution times, the data can be seen in \autoref{tbl:SSSP_execTimeComparison}. We show the results of Galois Push for the 5-node cluster in the table, since it is faster than Pull.

There are some, rather large differences in the execution times of each framework. Galois is consistently faster on a single computation node compared to the distributed setup. 
On both SSSP and BFS, the margin between the two setups is already noticeable. The distributed scenario requires from 2$\times$ (BFS, rMat28) to 7.8$\times$ (SSSP, wikipedia) the execution time of single-node Galois.
For PR however, the difference is multiple orders of magnitude large. Distributed Galois requires from 8.6$\times$ to 121$\times$ more time than single-node Galois.

Gemini's distributed calculation only requires more time on the smaller graphs (i.e. flickr to wikipedia). On the larger graphs, the added computation power can be leveraged to overcome the synchronization overhead of the distributed scenario. On rMat27, friendster and rMat28, Gemini is anywhere from 1.3$\times$ to 1.6$\times$ faster than on a single-node.
Giraph already uses the distributed cluster efficiently on the smaller graphs as well. Only on flickr is the distributed version slower than a single node. The differences for Giraph on the other graphs are rather large, single-node execution time is 2.25$\times$ and 3.7$\times$ the execution time of the distributed version on orkut and wikipedia respectively.

\input{sections/results/galoisSpeedup}