### 2020-09-19 documentation/paper/sections/overview/overview.tex
framework overview introduction

%!TEX root=../../main.tex

\section{Framework Overview}
The following paragraphs is a short overview describing functionality and characteristics of several state-of-the-art graph processing frameworks.
All of the described frameworks are part of our testing.
\todo

\subsection{Galois and Gluon}
\input{sections/overview/galois}

\subsection{Gemini}
\input{sections/overview/gemini}

\subsection{Giraph}
\input{sections/overview/giraph}

\subsection{Ligra}
\input{sections/overview/ligra}

\subsection{Polymer}
\input{sections/overview/polymer}

\input{sections/graphFormats}