### 2020-09-19 documentation/paper/main.tex
Ideen in Introduction rein gepastet

\documentclass[a4paper]{IEEEtran}

\usepackage[noadjust]{cite}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage{textcomp}
\usepackage{pifont}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{amssymb}

\usepackage{tabularx}
\def\tabularxcolumn#1{m{#1}} % vertical centering in tabularx cells
\newcolumntype{Y}{>{\centering\arraybackslash}X} % cell type Y for centered text with automatic line breaks
\newcommand*\rot{\rotatebox{45}} % rotates content 90deg ccw, mainly for table headers
\newcommand*\OK{\ding{51}} % a checkmark-style sign

\newcommand{\todo}[1]{{\color{red!90!black}//\bf{TODO:} #1}}

\bibliography{literature.bib}

\title{A Comparison of Graph Processing Systems}

\author{\IEEEauthorblockN{Simon König}
\IEEEauthorblockA{(3344789)
st156571@stud.uni-stuttgart.de}

\and

\IEEEauthorblockN{Leon Matzner}
\IEEEauthorblockA{(3315161)
st155698@stud.uni-stuttgart.de}

\and

\IEEEauthorblockN{Felix Rollbühler}
\IEEEauthorblockA{(3310069)
st154960@stud.uni-stuttgart.de}

\and

\IEEEauthorblockN{Jakob Schmid}
\IEEEauthorblockA{(3341630)
st157100@stud.uni-stuttgart.de}}

\date{\today}

\begin{document}

\maketitle

\begin{abstract}
\todo{Section}
\end{abstract}

\begin{IEEEkeywords}
graphs, distributed computing, Galois, Ligra, Polymer, Giraph, Gluon, Gemini
\end{IEEEkeywords}

\input{sections/introduction}

\input{sections/relatedWork}

\input{sections/preliminiaries}

\input{sections/overview/overview}

\section{Evaluation}
\todo{Einleitung}

\input{sections/testingMethods}

\input{sections/results/results}

\section{Discussion}
\todo{Hier zunächst mal ein kurzer Vergleich zwischen Distributed und Single Node wo sinnvoll, dann zwischen den Frameworks und so weiter..}

\section{Conclusion}
\todo{Section}

\clearpage

\section*{Acknowledgments}
We would like to thank our supervisor Heiko Geppert for his continued guidance and support.

\appendix
\label{app:installationGuides}\label{app:conversionTool}
We provide a set of installation guides, describing setup and usage of each of the presented frameworks. In case of Giraph, where for example no BFS algorithm is provided by the framework itself, the algorithm and a guide on how to include it in the framework is provided as well. Furthermore, the conversion tool we described to convert between the different graph input formats is available alongside the guides in our repository\footnote{\url{https://github.com/SerenGTI/Forschungsprojekt/tree/master/documentation}}.

% Beim Hinzufügen neuer Referenzen
% 1. die beiden Zeilen nach diesem Kommentar auskommentieren.
%    (gemeint ist \bibliographystyle... und \bibliography..)
%    Dann alle cache files im Ordner löschen.
% 2. Die folgenden Commands ausführen
% pdflatex main.tex
% bibtex main
% pdflatex main.tex
%    das letzte wirft dann Fehler aber das stimmt so..
% 3. Jetzt main.bbl öffnen und den Inhalt hier drunter rein kopieren.
%    Man kann kein \input benutzen, weil main.bbl leer ist, wenn die beiden
%    Zeilen auskommentiert sind. Aber wenn sie nicht auskommentiert sind
%    kann kein PDF erzeugt werden.. Was zur Hölle denken die sich.
% 4. Die folgenden beiden Zeilen auskommentieren und pdflatex
%    ausführen wie gewohnt.

% \bibliographystyle{IEEEtran}
% \bibliography{literature}

% Alles ab hier kommt aus main.bbl

% Generated by IEEEtran.bst, version: 1.12 (2007/01/11)
\begin{thebibliography}{10}
\providecommand{\url}[1]{#1}
\csname url@samestyle\endcsname
\providecommand{\newblock}{\relax}
\providecommand{\bibinfo}[2]{#2}
\providecommand{\BIBentrySTDinterwordspacing}{\spaceskip=0pt\relax}
\providecommand{\BIBentryALTinterwordstretchfactor}{4}
\providecommand{\BIBentryALTinterwordspacing}{\spaceskip=\fontdimen2\font plus
\BIBentryALTinterwordstretchfactor\fontdimen3\font minus
  \fontdimen4\font\relax}
\providecommand{\BIBforeignlanguage}[2]{{%
\expandafter\ifx\csname l@#1\endcsname\relax
\typeout{** WARNING: IEEEtran.bst: No hyphenation pattern has been}%
\typeout{** loaded for the language `#1'. Using the pattern for}%
\typeout{** the default language instead.}%
\else
\language=\csname l@#1\endcsname
\fi
#2}}
\providecommand{\BIBdecl}{\relax}
\BIBdecl

\bibitem{newman2010networks}
\BIBentryALTinterwordspacing
M.~Newman, \emph{Networks: An Introduction}.\hskip 1em plus 0.5em minus
  0.4em\relax OUP Oxford, 2010. [Online]. Available:
  \url{https://books.google.de/books?id=-DgTDAAAQBAJ}
\BIBentrySTDinterwordspacing

\bibitem{Polymer}
\BIBentryALTinterwordspacing
K.~Zhang, R.~Chen, and H.~Chen, ``Numa-aware graph-structured analytics,'' in
  \emph{Proceedings of the 20th ACM SIGPLAN Symposium on Principles and
  Practice of Parallel Programming}, ser. PPoPP 2015.\hskip 1em plus 0.5em
  minus 0.4em\relax New York, NY, USA: Association for Computing Machinery,
  2015, p. 183–193. [Online]. Available:
  \url{https://doi.org/10.1145/2688500.2688507}
\BIBentrySTDinterwordspacing

\bibitem{Ligra}
\BIBentryALTinterwordspacing
J.~Shun and G.~E. Blelloch, ``Ligra: A lightweight graph processing framework
  for shared memory,'' in \emph{Proceedings of the 18th ACM SIGPLAN Symposium
  on Principles and Practice of Parallel Programming}, ser. PPoPP ’13.\hskip
  1em plus 0.5em minus 0.4em\relax New York, NY, USA: Association for Computing
  Machinery, 2013, p. 135–146. [Online]. Available:
  \url{https://doi.org/10.1145/2442516.2442530}
\BIBentrySTDinterwordspacing

\bibitem{pregel}
\BIBentryALTinterwordspacing
G.~Malewicz, M.~H. Austern, A.~J. Bik, J.~C. Dehnert, I.~Horn, N.~Leiser, and
  G.~Czajkowski, ``Pregel: A system for large-scale graph processing,'' in
  \emph{Proceedings of the 2010 ACM SIGMOD International Conference on
  Management of Data}, ser. SIGMOD '10.\hskip 1em plus 0.5em minus 0.4em\relax
  New York, NY, USA: Association for Computing Machinery, 2010, p. 135–146.
  [Online]. Available: \url{https://doi.org/10.1145/1807167.1807184}
\BIBentrySTDinterwordspacing

\bibitem{pagerank}
\BIBentryALTinterwordspacing
S.~Brin and L.~Page, ``The anatomy of a large-scale hypertextual web search
  engine,'' \emph{Computer Networks and ISDN Systems}, vol.~30, no.~1, pp. 107
  -- 117, 1998, proceedings of the Seventh International World Wide Web
  Conference. [Online]. Available:
  \url{http://www.sciencedirect.com/science/article/pii/S016975529800110X}
\BIBentrySTDinterwordspacing

\bibitem{tao}
\BIBentryALTinterwordspacing
K.~Pingali, D.~Nguyen, M.~Kulkarni, M.~Burtscher, M.~A. Hassaan, R.~Kaleem,
  T.-H. Lee, A.~Lenharth, R.~Manevich, M.~M\'{e}ndez-Lojo, D.~Prountzos, and
  X.~Sui, ``The tao of parallelism in algorithms,'' in \emph{Proceedings of the
  32nd ACM SIGPLAN Conference on Programming Language Design and
  Implementation}, ser. PLDI '11.\hskip 1em plus 0.5em minus 0.4em\relax New
  York, NY, USA: Association for Computing Machinery, 2011, p. 12–25.
  [Online]. Available: \url{https://doi.org/10.1145/1993498.1993501}
\BIBentrySTDinterwordspacing

\bibitem{bsp}
\BIBentryALTinterwordspacing
L.~G. Valiant, ``A bridging model for parallel computation,'' \emph{Commun.
  ACM}, vol.~33, no.~8, p. 103–111, Aug. 1990. [Online]. Available:
  \url{https://doi.org/10.1145/79173.79181}
\BIBentrySTDinterwordspacing

\bibitem{congestModel}
\BIBentryALTinterwordspacing
S.~Forster and D.~Nanongkai, ``A faster distributed single-source shortest
  paths algorithm,'' \emph{2018 IEEE 59th Annual Symposium on Foundations of
  Computer Science (FOCS)}, Oct 2018. [Online]. Available:
  \url{http://dx.doi.org/10.1109/FOCS.2018.00071}
\BIBentrySTDinterwordspacing

\bibitem{Giraph}
\BIBentryALTinterwordspacing
{Apache Software Foundation}. (2020, Jun.) {Apache Giraph}. [Online].
  Available: \url{https://giraph.apache.org}
\BIBentrySTDinterwordspacing

\bibitem{graphx}
J.~E. Gonzalez, R.~S. Xin, A.~Dave, D.~Crankshaw, M.~J. Franklin, and
  I.~Stoica, ``Graphx: Graph processing in a distributed dataflow framework,''
  in \emph{Proceedings of the 11th USENIX Conference on Operating Systems
  Design and Implementation}, ser. OSDI'14.\hskip 1em plus 0.5em minus
  0.4em\relax USA: USENIX Association, 2014, p. 599–613.

\bibitem{powergraph}
\BIBentryALTinterwordspacing
J.~E. Gonzalez, Y.~Low, H.~Gu, D.~Bickson, and C.~Guestrin, ``Powergraph:
  Distributed graph-parallel computation on natural graphs,'' in \emph{10th
  {USENIX} Symposium on Operating Systems Design and Implementation ({OSDI}
  12)}.\hskip 1em plus 0.5em minus 0.4em\relax Hollywood, CA: {USENIX}
  Association, Oct. 2012, pp. 17--30. [Online]. Available:
  \url{https://www.usenix.org/conference/osdi12/technical-sessions/presentation/gonzalez}
\BIBentrySTDinterwordspacing

\bibitem{mapreduce}
J.~Dean and S.~Ghemawat, ``Mapreduce: Simplified data processing on large
  clusters,'' in \emph{In Proceedings of Operating Systems Design and
  Implementation (OSDI}, pp. 137--150.

\bibitem{hugepages}
\BIBentryALTinterwordspacing
A.~Panwar, A.~Prasad, and K.~Gopinath, ``Making huge pages actually useful,''
  in \emph{Proceedings of the Twenty-Third International Conference on
  Architectural Support for Programming Languages and Operating Systems}, ser.
  ASPLOS '18.\hskip 1em plus 0.5em minus 0.4em\relax New York, NY, USA:
  Association for Computing Machinery, 2018, p. 679–692. [Online]. Available:
  \url{https://doi.org/10.1145/3173162.3173203}
\BIBentrySTDinterwordspacing

\bibitem{origGalois}
\BIBentryALTinterwordspacing
M.~Kulkarni, K.~Pingali, B.~Walter, G.~Ramanarayanan, K.~Bala, and L.~P. Chew,
  ``Optimistic parallelism requires abstractions,'' in \emph{Proceedings of the
  28th ACM SIGPLAN Conference on Programming Language Design and
  Implementation}, ser. PLDI '07.\hskip 1em plus 0.5em minus 0.4em\relax New
  York, NY, USA: Association for Computing Machinery, 2007, p. 211–222.
  [Online]. Available: \url{https://doi.org/10.1145/1250734.1250759}
\BIBentrySTDinterwordspacing

\bibitem{Galois}
\BIBentryALTinterwordspacing
D.~Nguyen, A.~Lenharth, and K.~Pingali, ``A lightweight infrastructure for
  graph analytics,'' in \emph{Proceedings of the Twenty-Fourth ACM Symposium on
  Operating Systems Principles}, ser. SOSP ’13.\hskip 1em plus 0.5em minus
  0.4em\relax New York, NY, USA: Association for Computing Machinery, 2013, p.
  456–471. [Online]. Available: \url{https://doi.org/10.1145/2517349.2522739}
\BIBentrySTDinterwordspacing

\bibitem{vertGalois}
\BIBentryALTinterwordspacing
R.~Dathathri, G.~Gill, L.~Hoang, H.-V. Dang, A.~Brooks, N.~Dryden, M.~Snir, and
  K.~Pingali, ``Gluon: A communication-optimizing substrate for distributed
  heterogeneous graph analytics,'' in \emph{Proceedings of the 39th ACM SIGPLAN
  Conference on Programming Language Design and Implementation}, ser. PLDI
  2018.\hskip 1em plus 0.5em minus 0.4em\relax New York, NY, USA: Association
  for Computing Machinery, 2018, p. 752–768. [Online]. Available:
  \url{https://doi.org/10.1145/3192366.3192404}
\BIBentrySTDinterwordspacing

\bibitem{Gemini}
\BIBentryALTinterwordspacing
X.~Zhu, W.~Chen, W.~Zheng, and X.~Ma, ``Gemini: A computation-centric
  distributed graph processing system,'' in \emph{12th {USENIX} Symposium on
  Operating Systems Design and Implementation ({OSDI} 16)}.\hskip 1em plus
  0.5em minus 0.4em\relax Savannah, GA: {USENIX} Association, Nov. 2016, pp.
  301--316. [Online]. Available:
  \url{https://www.usenix.org/conference/osdi16/technical-sessions/presentation/zhu}
\BIBentrySTDinterwordspacing

\bibitem{scalability}
\BIBentryALTinterwordspacing
F.~McSherry, M.~Isard, and D.~G. Murray, ``Scalability! but at what {COST}?''
  in \emph{15th Workshop on Hot Topics in Operating Systems (HotOS
  {XV})}.\hskip 1em plus 0.5em minus 0.4em\relax Kartause Ittingen,
  Switzerland: {USENIX} Association, May 2015. [Online]. Available:
  \url{https://www.usenix.org/conference/hotos15/workshop-program/presentation/mcsherry}
\BIBentrySTDinterwordspacing

\bibitem{konect}
J.~Kunegis, ``Konect: the koblenz network collection,'' 05 2013, pp.
  1343--1350.

\bibitem{snap}
\BIBentryALTinterwordspacing
{Stanford University}. (2020, Sep.) Stanford network analysis project.
  [Online]. Available: \url{https://snap.stanford.edu}
\BIBentrySTDinterwordspacing

\bibitem{pbbs}
\BIBentryALTinterwordspacing
J.~Shun, G.~Blelloch, J.~Fineman, P.~Gibbons, A.~Kyrola, K.~Tangwonsan, and
  H.~V. Simhadri. (2020, Aug.) {Problem Based Benchmark Suite }. [Online].
  Available: \url{http://www.cs.cmu.edu/~pbbs/}
\BIBentrySTDinterwordspacing

\end{thebibliography}

\end{document}