### 2020-10-16 documentation/presentation/background.tex
dinge

\documentclass{meetings}

\usepackage{makecell}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{positioning}

\date{}

\usepackage[pages=all]{background}

\definecolor{left} {HTML}{001528}
\backgroundsetup{
scale=1,
color=black,
opacity=0.3,
angle=0,
contents={%
  \begin{tikzpicture}
  	\node [shading = axis,rectangle, left color=left, right color=white,shading angle=135, anchor=north, minimum width=\paperwidth, minimum height=\paperheight] (box) at (current page.north){};
  \end{tikzpicture}
  }%
}

\begin{document}
	\ 
\end{document}